---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, setup, include=FALSE}

source('requirements.R')

# load dataset
source('db.R')

# load functions
source('functions.R')

# get queries
source('queries.R')

```

```{r}
# Show first 5 entries of dataset
so_20 <- dbGetByQuery("SELECT * FROM survey_results_public LIMIT 5")

so_20_schema <- dbGetByQuery("SELECT * FROM survey_results_schema LIMIT 5")

so_20[1:5,]
```

# Initial Assessment
```{r}

#sprintf('The 2020 survey had %d respondents and %d variables.', dim(so_20)[1], dim(so_20)[2])

#head(so_20_schema)

```

```{r}
# Load world data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Create dataframe with all available countries
worldframe <- data.frame(x = world$name)

# Create dataframe with all survey countries
countries <- dbGetByCol(
  col = "Country"
)

countries <- countries %>% count(Country)

# Rename col for joining
countries <- countries %>%
  rename(
    x = Country
  )

# Join both dataframes together
world$surveyPop <- join(worldframe, countries, by='x')

# Change NA to 0
world$surveyPop[is.na(world$surveyPop)] <- 0
```

# Geography

```{r}
# Plot world map
worldpopmap <- ggplot(data=world) +
  # Set plot title
  ggtitle("World map", subtitle = paste0("(", length(unique(world$name)), " countries)")) +
  # Set x and y label
  xlab("Longitude") + ylab("Latitude") +
  # Fill worldmap with summarized survey data
  geom_sf(aes(fill = world$surveyPop$n)) +
  # Scale colors
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  # Customize legend title
  labs(fill = 'Frequency')

# display map
worldpopmap

```

```{r}
# Show previously plotted data in numbers

countries %>%
  arrange(-n)

```

# Developer Roles
About 55% of respondents identify as full-stack developers, and about 20% consider themselves mobile developers. The median number of developer type identifications per respondent this year is three, and the most common combinations include back-end, front-end, and full-stack developer. Pairs that are highly correlated include database administrator and system administrator, DevOps specialist and site reliability engineer, academic researcher and scientist, and designer and front-end developer.
```{r}

devTypes <- dbGetByCol(
  col = "DevType"
)

dT_split <- strsplit(devTypes$DevType, split=";")
dT_flatten <- unlist(dT_split)
dT <- data.frame(matrix(dT_flatten))

colnames(dT) <- colnames(devTypes)

dT %>% 
  ggplot( aes(x = fct_infreq(DevType))) + 
    geom_bar(fill='steelblue')+ 
    xlab("Dev Type") +
    coord_flip() +
    labs(title = "Developer Types",
         subtitle = paste0("All Respondents (", nrow(devTypes) ,")"),
         y = 'Count')
```


## Job Satisfactory
```{r}

sat <- dbGetByCol(
  col = "JobSat"
)

# Plot Overall Satisfactory
sat %>% count(JobSat) %>% arrange(n) %>% mutate(pct = prop.table(n), JobSat = fct_reorder(JobSat, n)) %>%
  ggplot(aes(
    x=JobSat, 
    y=n,
    label= scales::percent(pct))) +
    geom_col(position = 'dodge', fill = 'steelblue') + 
    geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = 0.5,    # nudge above top of bar
              hjust = -0.1,
              size = 3) + 
    xlab("Level of Job Satisfactory") +
    ylab("Amount of Respondents") +
    labs(title = "Overall Satisfactory",
         subtitle = paste0("All Respondents (", nrow(sat) ,")")) +
    coord_flip()

```

```{r fig.height=7, fig.width=14}

# Plot Satisfactory dependent with educational attainment

satByEdu <- dbGetByQuery(
  query = "SELECT count(JobSat) as cJobSat,JobSat,EdLevel
FROM survey_results_public
WHERE JobSat IS NOT NULL
AND EdLevel IS NOT NULL
GROUP BY JobSat, EdLevel
ORDER BY cJobSat DESC;" 
)

satByEdu[,"cJobSat"] <- as.integer(satByEdu[,"cJobSat"])

satByEdu %>% 
  ggplot(aes(
    x = JobSat,
    y = cJobSat,
    fill = EdLevel
  )) + geom_bar(stat = "identity") + 
  theme(legend.position="bottom", 
        legend.box="vertical", 
        legend.margin=margin()
        ) +
  ylab("Count") + xlab("Level of Satisfaction") +
  labs(
    fill = "Educational Attainment",
    title = "Overall Satisfactory by Educational Attainment",
    subtitle = paste0("All Respondents (", sum(satByEdu$cJobSat) ,")")
  ) + coord_flip() + guides(fill=guide_legend(nrow=3,byrow=TRUE))

```




## Coding as a hobby

Many developers work on code outside of work. About 78% of our respondents say that they code as a hobby. Other responsibilities outside of software can reduce developers' engagement in coding as a hobby; developers who say they have children or other caretaking responsibilities are less likely to code as a hobby. Respondents who are women are also less likely to say they code as a hobby.

```{r}
hobby <- dbGetByQuery("SELECT 
Hobbyist, 
count(Hobbyist) as CountHobbyist, 
Gender 
FROM new_schema.survey_results_public
WHERE
Gender = 'Woman' or
Gender = 'Man'
group by 
Hobbyist,
Gender;
")

hobby[,"CountHobbyist"] <- as.integer(hobby[,"CountHobbyist"])

hobby <- hobby %>% mutate(pct = prop.table(CountHobbyist))

# Create reusable plot without data
gg_hobby <- ggplot(data = data.frame(), aes(x = Hobbyist, 
                                            y = CountHobbyist, 
                                            fill=Gender,
                                            label= scales::percent(pct))) +
  geom_col(position = 'dodge') + 
  geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = -0.5,    # nudge above top of bar
              size = 3) + 
  labs(x = 'Hobbyist', 
       y = NULL, 
       fill= 'Gender', 
       title= 'Amount of Hobbyists by Gender',
       subtitle= paste0("All Respondents (", sum(hobby$CountHobbyist) ,")"))

# Add data after
gg_hobby %+%
  hobby


```

# Experience
## Years since Learning to Code

There is a wide range of experience among developers who visit Stack Overflow, from seasoned developers who learned to code more than 30 years ago (approximately 15%), to a sizable percentage of developers (17%) who learned how to code less than five years ago. Of the professional developers on Stack Overflow, approximately 40% learned to code less than 10 years ago.

```{r}

exp <- dbGetByCol(
  col = "YearsCode"
)



# Create new dataframe for storage
newExp <- data.frame()
  
# Convert data to a numerical column
exp$nYearsCode <- as.integer(exp$YearsCode)
exp[grep('Less', exp$YearsCode),]$nYearsCode <- 0
exp[grep('More', exp$YearsCode),]$nYearsCode <- 50

# Select data relted for plotting

newExp['Less than 5 years', 'n'] <- (exp %>% filter(nYearsCode < 5))$n %>% sum() 
newExp['5 to 9 years', 'n'] <- (exp %>% filter(between(nYearsCode, 5, 9)))$n %>% sum()
newExp['10 to 14 years', 'n'] <- (exp %>% filter(between(nYearsCode, 10, 14)))$n %>% sum()
newExp['15 to 19 years', 'n'] <- (exp %>% filter(between(nYearsCode, 15, 19)))$n %>% sum()
newExp['20 to 24 years', 'n'] <- (exp %>% filter(between(nYearsCode, 20, 24)))$n %>% sum()
newExp['25 to 29 years', 'n'] <- (exp %>% filter(between(nYearsCode, 25, 29)))$n %>% sum()
newExp['30 to 34 years', 'n'] <- (exp %>% filter(between(nYearsCode, 30, 34)))$n %>% sum()
newExp['35 to 39 years', 'n'] <- (exp %>% filter(between(nYearsCode, 35, 39)))$n %>% sum()
newExp['40 to 44 years', 'n'] <- (exp %>% filter(between(nYearsCode, 40, 45)))$n %>% sum()
newExp['45 to 49 years', 'n'] <- (exp %>% filter(between(nYearsCode, 45, 49)))$n %>% sum()
newExp['50 years or more', 'n'] <- (exp %>% filter(nYearsCode >= 50))$n %>% sum() 

#
#between_c(exp, exp$YearsCode, 5, 9)$n

# Compute percentages
newExp$perc <- perc(newExp$n, sum(exp$n))

# Move idx names to a col
newExp$desc <- rownames(newExp)

newExp %>%
  mutate(desc = fct_reorder(desc, desc(-n))) %>%
  ggplot(aes(x= desc, y= perc)) + 
    geom_bar(stat='identity', fill='steelblue') +
    xlab("") +
    coord_flip() + 
    labs(y = 'Percentage', 
         title='Years since learning to Code',
         subtitle = paste0("All Respondents (", nrow(exp) ,")"))

```

## Writing That First Line of Code
Of all of the respondents, over 54% wrote their first line of code, whether it was a web page or a hello world program, by the age of 16. People who wrote their first line of code in their 20s accounted for 13% of the respondents. When looking at the average age by country, respondents from countries such as Brazil and India tend to start writing code a full two years later compared to developers in countries such as Poland and Germany, who on average start coding by the age of 15.

```{r}
fcode <- dbGetByCol(
  col = "Age1stCode"
)

newFCode <- data.frame()

# Convert data to a numerical column
fcode$nAge1stCode <- as.integer(fcode$Age1stCode)

# Make string col readable
fcode[grep('Younger', fcode$Age1stCode),]$nAge1stCode <- 10
fcode[grep('Older', fcode$Age1stCode),]$nAge1stCode <- 30

# Select data related for plotting

newFCode['Younger than 10 years', 'n'] <- (fcode %>% filter(nAge1stCode < 10))$n %>% sum()
newFCode['10 to 11 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 10, 11)))$n %>% sum()
newFCode['12 to 13 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 12, 13)))$n %>% sum()
newFCode['14 to 15 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 14, 15)))$n %>% sum()
newFCode['16 to 17 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 16, 17)))$n %>% sum()
newFCode['18 to 19 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 18, 19)))$n %>% sum()
newFCode['20 to 21 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 20, 21)))$n %>% sum()
newFCode['22 to 23 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 22, 23)))$n %>% sum()
newFCode['24 to 25 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 24, 25)))$n %>% sum()
newFCode['26 to 27 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 26, 27)))$n %>% sum()
newFCode['28 to 29 years old', 'n'] <- (fcode %>% filter(between(nAge1stCode, 28, 29)))$n %>% sum()
newFCode['30 years old or older', 'n'] <- (fcode %>% filter(nAge1stCode >= 30))$n %>% sum() 

# Print sum of rows to check if all rows have been added
# newFCode %>% sum()

# Compute percentages
newFCode$perc <- perc(newFCode$n, sum(fcode$n))

# Move idx names to a col
newFCode$desc <- rownames(newFCode)

newFCode %>%
  mutate(desc = fct_reorder(desc, desc(-n))) %>%
  ggplot( aes(x= desc, y= perc)) + 
    geom_bar(stat='identity', fill='steelblue') +
    xlab("") +
    coord_flip() + 
    labs(y = 'Percentage', title='Writing That First Line of Code',
         subtitle = paste0("All Respondents (", nrow(fcode) ,")"))
```
```{r}

fcode_country <- dbGetByQuery("SELECT avg(Age1stCode) as 'AvgAge1stCode', Country, Gender
FROM survey_results_public
WHERE CONCAT(Country, Gender) IS NOT NULL
AND Gender = 'Woman' OR Gender = 'Man'
GROUP BY Country, Gender ORDER BY Country;");

LOWER_BOUND <- 10
UPPER_BOUND <- 12

# Handle UTF8TOWCS ERROR
fcode_country$Country <- str_replace_all(fcode_country$Country, "[^[:graph:]]", " ")

fcode_country <- fcode_country %>% filter(between(AvgAge1stCode, LOWER_BOUND, UPPER_BOUND))

fcode_country %>% 
  ggplot(
    aes(
      x = AvgAge1stCode, 
      y = Country, 
      fill = Gender)) + 
  geom_col(position = 'dodge') + 
  geom_text(
    label=sprintf("%.1f", round(fcode_country$AvgAge1stCode, digits = 2)),
    position = position_dodge(0.9), hjust = -0.2
    ) +
  labs(
    y = 'Country', 
    x = 'Average Age - First Line of Code',
    title='Writing That First Line of Code',
    subtitle = paste0('by Country (Mean Age between ', LOWER_BOUND, ' and ', UPPER_BOUND, ')')
  ) + xlim(0, 15)

```


# Education
## Educational Attainment
Approximately 75% of respondents worldwide completed at least the equivalent of a bachelor's degree or higher. This is consistent with what we've seen in previous years.

```{r}

edu <- dbGetByCol(
  col = "EdLevel"
)

edu %>% 
  count(EdLevel) %>%
  mutate(EdLevel = fct_reorder(EdLevel, desc(-n))) %>%
    ggplot( aes(x=EdLevel, y=n)) + 
      geom_bar(stat='identity', fill='steelblue') +
      xlab("") +
      coord_flip() +
      labs(y = 'Cumulative Number', title= 'Educational Attainment',
           subtitle = paste0("All Respondents (", nrow(edu) ,")"))

```
# Salary

```{r}
comp_country <- dbGetByQuery("SELECT CompFreq, CompTotal, Gender
FROM survey_results_public
WHERE Country = 'Germany'
AND CONCAT (CompFreq, CompTotal, Gender) IS NOT NULL;")

# Group genders
comp_country$Gender <- apply(comp_country,1,genderApply)

# Convert to integer
#comp_country$CompTotal <- as.integer(comp_country$CompTotal)

# Filter by yearly payed compensation
yearly <- comp_country %>% filter(CompFreq == "Yearly") 

# Divide yearly payed comp by 12
yearly$CompTotal <- as.numeric(as.character(yearly$CompTotal)) / 12

# Concat again
comp_country <- rbind(comp_country %>% filter(CompFreq == "Monthly"), yearly)

comp_country <- comp_country[which(comp_country$CompTotal < 15000),]

comp_country %>%
  ggplot(
    aes(x = CompTotal, fill=Gender)
  ) + 
  geom_histogram(
    aes(y=..density..),
    binwidth=500,
    alpha=.5, position = 'identity'
    ) + 
  geom_density(alpha=.5) +
  facet_grid(Gender ~ .) + 
  labs(
    title="Distribution of Salary in Germany",
    subtitle = paste0("All Respondents (", nrow(comp_country) ,")"),
    x="Monthly payed Salary",
    y="Density"
    ) +
  scale_y_continuous()

```

```{r fig.width = 14, fig.height=10}

salaryByDevType <- dbGetByQuery(query = 'SELECT DevType, CompTotal, CompFreq, YearsCodePro
FROM survey_results_public
WHERE (CompTotal < 150000 AND CompFreq = "Yearly")
OR (CompTotal < 15000 AND CompFreq = "Monthly")
AND CONCAT(DevType, CompTotal, CompFreq) IS NOT NULL;')

salaryByDevType <- salaryByDevType %>% na.omit()

# Convert less than 1 year values
salaryByDevType[rownames(salaryByDevType[which(salaryByDevType$YearsCodePro == "Less than 1 year"),]), "YearsCodePro"] <- "1"
# Convert more than 50 years values
salaryByDevType[rownames(salaryByDevType[which(salaryByDevType$YearsCodePro == "More than 50 years"),]), "YearsCodePro"] <- "50"

salaryByDevType$YearsCodePro <- as.integer(salaryByDevType$YearsCodePro)

resp <- nrow(salaryByDevType)

# Split into monthly and yearly
monthly <- salaryByDevType %>% filter(CompFreq == "Monthly")
yearly <- salaryByDevType %>% filter(CompFreq == "Yearly")

# Compute yearly salary
monthly$CompTotal <- monthly$CompTotal * 12

# Combine them together
salaryByDevType <- rbind(monthly, yearly)

s <- strsplit(salaryByDevType$DevType, split=";")

# Change dimensions of dataframe
salaryByDevType <- data.frame(
  CompTotal = rep(salaryByDevType$CompTotal, sapply(s, length)),
  YearsProCode = rep(salaryByDevType$YearsCodePro, sapply(s, length)),
  DevType = unlist(s)
)

# Outliers
Q <- quantile(salaryByDevType$CompTotal, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(salaryByDevType$CompTotal)

up <-  Q[2]+2.5*iqr # Upper Range  
low<- Q[1]-2.5*iqr # Lower Range

salaryByDevType <- subset(salaryByDevType, salaryByDevType$CompTotal > (Q[1] - 1.5*iqr) & salaryByDevType$CompTotal < (Q[2]+1.5*iqr))

# Arrange Data
types <- salaryByDevType %>% count(DevType)

salaryByDevType <- salaryByDevType %>% group_by(DevType) %>% summarise(
  CompTotal = mean(CompTotal),
  YearsProCode = mean(YearsProCode)
)

# Merge both frames by DevType (Ensures that the correct count is in the appropriate row)
salaryByDevType <- merge(salaryByDevType, types, by=c("DevType"))

# Outliers of Count / n
Q <- quantile(salaryByDevType$n, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(salaryByDevType$n)

up <-  Q[2]+2.5*iqr # Upper Range  
low<- Q[1]-2.5*iqr # Lower Range

salaryByDevType <- subset(salaryByDevType, salaryByDevType$n > (Q[1] - 1.5*iqr) & salaryByDevType$n < (Q[2]+1.5*iqr))

ggplot(salaryByDevType, 
       aes(
         x = YearsProCode,
         y = CompTotal,
         size = n,
         color = abs(n)
       )) +
  geom_point(
    alpha = .7
  ) +
  geom_text_repel(
    aes(
      label = DevType
    ),
    color = "black", 
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) + 
  scale_size(range = c(.1, 10), name="Number of respondents") +
  scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
  scale_colour_continuous(guide = FALSE) +
  labs(
    title = "Salary and Experience by Developer Type",
    subtitle = paste0("All Respondents (", resp ,")"),
    x = "Average years of professional programming experience",
    y = "Median salary (USD)"
  ) +
  scale_y_continuous(labels = scales::dollar) +
  geom_smooth(
    #method=lm,
    method=loess,
    formula = y ~ x,
    alpha=.2,
    )




```


# Salary and Experience by Language

```{r fig.width=12, fig.height=10}

salExp <- dbGetByQuery('SELECT CompFreq, CompTotal, YearsCodePro, LanguageWorkedWith
FROM survey_results_public
WHERE (CompTotal < 150000 AND CompFreq = "Yearly")
OR (CompTotal < 15000 AND CompFreq = "Monthly")
AND CONCAT(CompFreq, CompTotal, LanguageWorkedWith, YearsCodePro) IS NOT NULL;')

salExp <- salExp %>% na.omit()

# Convert less than 1 year values
salExp[rownames(salExp[which(salExp$YearsCodePro == "Less than 1 year"),]), "YearsCodePro"] <- "1"
# Convert more than 50 years values
salExp[rownames(salExp[which(salExp$YearsCodePro == "More than 50 years"),]), "YearsCodePro"] <- "50"

salExp$YearsCodePro <- as.integer(salExp$YearsCodePro)

resp <- nrow(salExp)

# Split into monthly and yearly
monthly <- salExp %>% filter(CompFreq == "Monthly")
yearly <- salExp %>% filter(CompFreq == "Yearly")

# Compute yearly salary
monthly$CompTotal <- monthly$CompTotal * 12

# Combine them together
salExp <- rbind(monthly, yearly)

s <- strsplit(salExp$LanguageWorkedWith, split = ";")

# Change dimensions of dataframe
salExp <- data.frame(
              CompTotal = rep(salExp$CompTotal, sapply(s, length)),
              YearsProCode = rep(salExp$YearsCodePro, sapply(s, length)),
              LanguageWorkedWith = unlist(s)
            )

# Outliers
Q <- quantile(salExp$CompTotal, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(salExp$CompTotal)

up <-  Q[2]+2.5*iqr # Upper Range  
low<- Q[1]-2.5*iqr # Lower Range

salExp<- subset(salExp, salExp$CompTotal > (Q[1] - 1.5*iqr) & salExp$CompTotal < (Q[2]+1.5*iqr))

langs <- (salExp %>% count(LanguageWorkedWith))$n

salExp <- salExp %>% group_by(LanguageWorkedWith) %>% summarise(
  CompTotal = mean(CompTotal), 
  YearsProCode = mean(YearsProCode)
  )

salExp["Respondents"] <- langs

ggplot(salExp,
    aes(
      x = YearsProCode,
      y = CompTotal,
      size = Respondents,
      color = abs(Respondents)
    )
  ) + 
  geom_point(
    alpha = .7
  ) +
  geom_text_repel(
    aes(
      label=LanguageWorkedWith
    ),
    fill = "black", 
    color = "black", 
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) + 
  scale_size(range = c(.1, 10), name="Number of respondents") +
  scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
  scale_colour_continuous(guide = FALSE) +
  labs(
    title = "Salary and Experience by Language",
    subtitle = paste0("All Respondents (", resp ,") - Professional Coding Experience between 6 and 15 years"),
    x = "Average years of professional programming experience",
    y = "Median salary (USD)"
  ) +
  scale_y_continuous(labels = scales::dollar) +
  geom_smooth(
    method=loess,
    formula = y ~ x,
    alpha=.1)

```
